Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib

### 2023-09-20 texor.Rmd
two more issues completed

---
title: "Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib"
date: "`r Sys.Date()`"
abstract: >
  In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX. The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers. A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output. This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Abhishek Ulayil
    affiliation: Institute of Actuaries of India (student)
    address:
    - Mumbai, India
    orcid: 0009-0000-6935-8690
    email: perricoq@outlook.com
  - name: Mitch O'Hara-Wild
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0001-6729-7695
    email: mail@mitchelloharawild.com
  - name: Christophe Dervieux
    affiliation: Posit PBC
    address:
    - Paris, France
    orcid: 0000-0003-4474-2498
    email: christophe.dervieux@gmail.com
  - name: Heather Turner
    affiliation: University of Warwick
    address:
    - Coventry, United Kingdom
    orcid: 0000-0002-1256-3375
    email: ht@heatherturner.net
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

@@

1. Articles can include interactive graphics and tables. 
2. The format is more accessible to screen readers and other assistive technologies, making the work more accessible to vision-impaired researchers.
3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
4. Search engines can easily access the full text of articles, facilitating discovery of published articles.

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element. In an R Markdown code chunk generating a figure, this can be set using `fig.alt`. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

>>>>

---
title: "Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib"
date: "`r Sys.Date()`"
abstract: >
  In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX. The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print disabled readers. A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output. This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Abhishek Ulayil
    affiliation: Institute of Actuaries of India (student)
    address:
    - Mumbai, India
    orcid: 0009-0000-6935-8690
    email: perricoq@outlook.com
  - name: Mitch O'Hara-Wild
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0001-6729-7695
    email: mail@mitchelloharawild.com
  - name: Christophe Dervieux
    affiliation: Posit PBC
    address:
    - Paris, France
    orcid: 0000-0003-4474-2498
    email: christophe.dervieux@gmail.com
  - name: Heather Turner
    affiliation: University of Warwick
    address:
    - Coventry, United Kingdom
    orcid: 0000-0002-1256-3375
    email: ht@heatherturner.net
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Melbourne, Australia
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

@@

1. Articles can include interactive graphics and tables. 
2. The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers. (See @W3C for current guidelines.)
3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
4. Search engines can easily access the full text of articles, facilitating discovery of published articles.

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element. In an R Markdown code chunk generating a figure, this can be set using `fig.alt`. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

### 2023-09-20 texor.tex
two more issues completed

\abstract{%
In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX. The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers. A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output. This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Articles can include interactive graphics and tables.
\item
  The format is more accessible to screen readers and other assistive technologies, making the work more accessible to vision-impaired researchers.
\item
  HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
\item
  Search engines can easily access the full text of articles, facilitating discovery of published articles.
\end{enumerate}

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element. In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

There are many other journals that have a similar backlog of legacy articles. A current and ongoing project is being conducted by {``HTML as an Accessible Format for Papers''} (2023) to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML ({``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter''} 2023). The work conducted for the R Journal might of interest to similar small journals that operate without paid staff. Converting legacy articles provides accessibility to a wealth of educational material.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-arxiv}{}}%
{``HTML as an Accessible Format for Papers.''} 2023. https://info.arxiv.org/about/accessible\_HTML.html.

@@

\leavevmode\vadjust pre{\hypertarget{ref-latexML}{}}%
{``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter.''} 2023. https://math.nist.gov/~BMiller/LaTeXML/.

@@

@@

>>>>

\abstract{%
In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX. The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print disabled readers. A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output. This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Articles can include interactive graphics and tables.
\item
  The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers. (See Sheri Byrne-Haber (2023) for current guidelines.)
\item
  HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
\item
  Search engines can easily access the full text of articles, facilitating discovery of published articles.
\end{enumerate}

@@

There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community. Authors can become more aware of good practice. For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element. In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}. Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines. It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

There are many other journals that have a similar backlog of legacy articles. A current and ongoing project is being conducted by arXiv Accessibility Project (2023) to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML (Miller and Ginev 2023). The work conducted for the R Journal might of interest to similar small journals that operate without paid staff. Converting legacy articles provides accessibility to a wealth of educational material.

@@

\leavevmode\vadjust pre{\hypertarget{ref-arxiv}{}}%
arXiv Accessibility Project. 2023. {``HTML as an Accessible Format for Papers.''} \url{https://info.arxiv.org/about/accessible_HTML.html}.

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-latexML}{}}%
Miller, Bruce, and Deyan Ginev. 2023. {``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter.''} \url{https://math.nist.gov/~BMiller/LaTeXML/}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-W3C}{}}%
Sheri Byrne-Haber, Charles LaPierre, David Fazio. 2023. {``W3C Accessibility Maturity Model.''} \url{https://www.w3.org/TR/maturity-model/}.
